ICS_Cytokine_Analysis

### 2019-05-29 Cytokine_Analysis.Rmd
cleaning up to source functions

---
title: "Cytokine_Analysis"
author: "Taryn McLaughlin"
date: "9/21/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 10.5, fig.width = 7.5)

library(dplyr)
library(reshape2)
Cytokine<-read.csv("/Applications/Old Computer/Day Lab/Data_Flow/Clean/Background Subtracted Cytokine.csv")
Cytokine <- mutate(Cytokine, G_plus = G_4_13_x + G_4_x_x + G_x_13_x + G_x_x_x) %>%
    mutate(T_plus = x_4_13_T + x_4_x_T + x_x_13_T + x_x_x_T) %>%
    mutate(G_T_plus = G_4_x_T + G_x_13_T + G_x_x_T)%>%
    mutate(IL4_plus = G_4_x_T + G_4_x_x + x_4_x_T + x_4_x_x) %>%
    mutate(IL13_plus = G_x_13_T + G_x_13_x + x_x_13_T + x_x_13_x) %>%
    mutate(IL4_IL13_plus = G_4_13_x + x_4_13_T + x_4_13_x)
Cytokine$TB<-factor(Cytokine$TB, levels=c("N","HC","LTBI","TB"))
Cytokine$SM<-factor(Cytokine$SM, levels=c("N","X","SM"))
```

#Statistical Analysis
I wrote a function called pvals which evaluates the relevant two sample two-tailed  p-values using the non-parametric Mann Whitney test (the command is wilcox.test in R). The function then corrects for multiple comparisons in two ways. The first uses  a Bonferonni correction to control the Family Wise Error Rate. The second controls the False Discovery Rate, the expected proportion of false discoveries amongst the rejected hypotheses. The false discovery rate is a less stringent condition than the family-wise error rate, so these methods are more powerful than the others. The function is below if you would like to see the details.

```{r p-value function, echo=TRUE}
pvals<-function(data, column){

    data<-filter(data, SM!="N")
    data$SM<-factor(data$SM, levels=c("X","SM"))
    data$TB<-factor(data$TB, levels=c("HC","LTBI","TB"))

    x<-split(data,data$TB) 
    A<-lapply(x, function(g) wilcox.test(g[,column]~g[,"SM"]))

    y<-split(data,data$SM) 
    B<-lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "LTBI")))
    C<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("LTBI", "TB")))
    D<- lapply(y, function(g) 
        wilcox.test((g[,column]~g[,"TB"]), subset=g$TB %in% c("HC", "TB")))

    labels<-c("HC SM+ to HC SM-", "LTBI SM+ to LTBI SM-", "TB SM+ to TB SM-",
              "HC SM+ to LTBI SM+", "LTBI SM+ to TB SM+", "HC SM+ to TB SM+",
              "HC SM- to LTBI SM-", "LTBI SM- to TB SM-", "HC SM- to TB SM-"
    )
    pvals<-c(A$HC$p.value, A$LTBI$p.value, A$TB$p.value,
             B$SM$p.value, C$SM$p.value,  D$SM$p.value,
             B$X$p.value,  C$X$p.value,   D$X$p.value)

    bonf_adj_pvals<-round(p.adjust(pvals, method="bonferroni"), 4)

    fdr_adj_pvals<-round(p.adjust(pvals, method="fdr"), 4)

    table<-(cbind(labels, round(pvals, 4), bonf_adj_pvals, fdr_adj_pvals))
    colnames(table)<-c("Comparison", paste(column,"p-values"," "), 
                       "Bonferonni Corrected p-values", "FDR Corrected p-vales")
    table
}
```

#Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it incredibly cluttered looking, however I did allow outliers to display.

```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
Boolean<-select(Cytokine, Donor, TB, SM, Stim, G_4_13_T, G_4_x_T, G_x_13_T, G_4_13_x, x_4_13_T, G_4_x_x, G_x_13_x,
                x_4_x_T, x_x_13_T, G_x_x_T , G_x_x_x, x_x_x_T,  x_4_13_x, x_4_x_x, x_x_13_x)

melt<-melt(Boolean,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_4_13_T", "G_4_x_T", "G_x_13_T", "G_4_13_x", "x_4_13_T","G_4_x_x", 
                                           "G_x_13_x","x_4_x_T","x_x_13_T", "G_x_x_T" , "G_x_x_x","x_x_x_T",  
                                           "x_4_13_x","x_4_x_x", "x_x_13_x"))
melt<-filter(melt, !is.na(label))

#all the boolean plots
g<-ggplot(melt, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(Stim~TB, scale="free")+
    labs(title="Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

#subset by stim
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="PMA Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="WCL Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");data<-filter(melt, Stim=="PEP")

data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="PEP Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="SEA")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="SEA Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");

data<-filter(melt, Stim=="SWAP")
g<-ggplot(data, aes(x=label, y=value, fill=TB, alpha=SM))
g + geom_boxplot(size=1, position=position_dodge(width = 1)) + 
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=90, vjust=0.6)) + 
    facet_grid(TB~., scale="free")+
    labs(title="SWAP Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="Cytokine Subsets",
         y="Frequency of Cytokine+ CD4+ T cells (%)");
```

#T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups.

```{r TH, echo=FALSE, fig.cap="Breakdown of Cytokine Definitions"}
knitr::include_graphics("/Applications/Old Computer/Day Lab/Data_Flow/TH_Group_Explanation.png")
```

```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
TH<-select(Cytokine, Donor, TB, SM, Stim, Cyt.Freq, TH1.Freq, TH1.2.Freq, TH2.Freq)

melt<-melt(TH,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("Cyt.Freq", "TH1.Freq", "TH1.2.Freq", "TH2.Freq"))
melt<-filter(melt, !is.na(label))

##### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PMA")
kable(pvals(data, "TH1.Freq"), caption = "Non Specific CD4+ T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

##### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="WCL")
kable(pvals(data, "TH1.Freq"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

##### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(TH, Stim=="PEP")
kable(pvals(data, "TH1.Freq"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(data, "TH2.Freq"))
kable(pvals(data, "TH1.2.Freq"))

```
\newpage

@@

```{r Individual Cytokines but not single producers}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
cyt<-select(Cytokine, Donor, TB, SM, Stim, IFNg, TNFa, IL4, IL13)

melt<-melt(cyt,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("IFNg", "TNFa", "IL4", "IL13"))
melt<-filter(melt, !is.na(label))

#### PMA
data<-filter(melt, Stim=="PMA")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PMA")
kable(pvals(data, "IFNg"), caption = "Non Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

#### WCL
data<-filter(melt, Stim=="WCL")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="WCL")
kable(pvals(data, "IFNg"), caption = "Mtb Whole Cell Lysate Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

#### Peptide
data<-filter(melt, Stim=="PEP")
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Peptide Specific CD4+ T cell Responses", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(cyt, Stim=="PEP")
kable(pvals(data, "IFNg"), caption = "Mtb Peptide Specific CD4 T cell Responses")
kable(pvals(data, "TNFa"))
kable(pvals(data, "IL4"))
kable(pvals(data, "IL13"))

#True Single and Double Cytokine Positive CD4+ T Cells
These are the frequencies of cells truly ONLY producing:  
1. IFNg or TNFa or the combination of both AND NOT IL4 or IL13  
2. IL-4 or IL-13 or the combination of both AND NOT IFNg or TNFa  
These are the subgroups within the grouped TH1 vs TH2 lineage analysis. They are essentially the last 6 groups of the boolean data.

```{r Single and Double Producers Lin Spec graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
SP_lin<-rename(Cytokine, G_single =  G_x_x_x, T_single = x_x_x_T, IL4_single = x_4_x_x, IL13_single = x_x_13_x, 
               G_T = G_x_x_T, IL4_IL13 = x_4_13_x) %>%
    select(Donor, TB, SM, Stim, G_single, T_single, IL4_single, IL13_single, G_T, IL4_IL13)

melt<-melt(SP_lin,id.vars=c("TB","SM", "Stim"))
melt$value<-as.numeric(melt$value)
melt<-filter(melt, !(is.na(value)))
melt<-filter(melt, !(value<0))
melt$label = factor(melt$variable,levels=c("G_single", "T_single", "G_T", "IL4_single", "IL13_single", "IL4_IL13"))
melt<-filter(melt, !is.na(label))

#### PMA
# Th1 plots
data<-filter(melt, Stim=="PMA" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="PMA" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Non-Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="PMA")
kable(pvals(data, "G_single"), caption = "Non-Specific CD4+ T cell Responses")
kable(pvals(data, "T_single")) 
kable(pvals(data,"G_T"))
kable(pvals(data, "IL4_single"))
kable(pvals(data, "IL13_single")) 
kable(pvals(data, "IL4_IL13"))

#### WCL
# Th1 plots
data<-filter(melt, Stim=="WCL" & label %in% c("G_single", "T_single", "G_T"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")
#TH2 plots
data<-filter(melt, Stim=="WCL" & label %in% c("IL4_single", "IL13_single", "IL4_IL13"))
g<-ggplot(data, aes(x=SM, y=value, fill=TB, alpha=SM))
g + geom_boxplot(outlier.shape = NA, size=1) + 
    geom_jitter(width=.1,height=0, shape=1,size=2) +
    scale_fill_manual(values =c("#e0e0e0", "#1a9850" , "#2166ac", "#b2182b"))+
    scale_alpha_manual(values=c(0.5,0.5,1))+
    scale_x_discrete(breaks=c("N","X","SM"),
                     labels=c("Naive", "SM(-)", "SM(+)"))+
    theme_bw()+ theme(legend.position="none")+
    theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    facet_grid(label~TB, scale="free")+
    labs(title="Mtb Whole Cell Lysate Specific CD4+ Cytokine Frequencies", 
         subtitle="",
         caption="",
         x="TB and S. mansoni Status",
         y="Frequency of Cytokine+ CD4+ T cells (%)")

data<-filter(SP_lin, Stim=="WCL")
kable(pvals(data, "G_single"), caption = "Mtb Whole Cell Lysate Specific CD4+ T cell Responses")
kable(pvals(data, "T_single")) 
kable(pvals(data,"G_T"))
kable(pvals(data, "IL4_single"))
kable(pvals(data, "IL13_single")) 
kable(pvals(data, "IL4_IL13"))

>>>>

---
title: "ICS_Cytokine_Analysis"
author: "Taryn McLaughlin"
date: "9/21/2018"
output: 
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE, comment=NA)
library(dplyr)
library(knitr)
library(ggplot2)
library(gridExtra)
library(GGally)
library(ggpubr)
library(cowplot)
```

```{r source}
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/tryna function.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/pvals_TB-SM.R")
source("/Users/tarynam/Desktop/Miscellaneous-Scripts/plot_functions.R")
```

```{r load data, echo=FALSE}
setwd("/Applications/Old Computer/Day Lab/Flow-Data/clean/")

CD4<-plot_filter(read.csv("ICS_CD4_Cytokine_clean.csv"))

CD8<-plot_filter(read.csv("ICS_CD8_Cytokine_clean.csv"))

MFI<-plot_filter(read.csv("ICS_CD4_MFI_clean.csv"))

Compass<-plot_filter(read.csv("ICS_Compass_scores_complete.csv"))

GD<-plot_filter(read.csv("ICS_GD_Cytokine_clean.csv"))

data<-read.csv("ICS_Donor_Info_clean.csv")
data$TB<-gsub("BCG|Unknown", "Naïve", data$TB)
data$TB<-factor(data$TB, levels=c("Naïve","Healthy Control","LTBI","Active TB"))
data$SM<-gsub("US|Not US|Unknown", "Naïve", data$SM)
data$SM<-factor(data$SM, levels=c("Naïve","Negative","Schisto+"))
Donor<-data
```
#P-values for Cytokine Comparisons

##PMA & Ionomycin
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PMA") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

##Whole Cell Lysate
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="WCL") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

##CFP-10 & ESAT-6
```{r}
library(knitr)
#creates a list of character strings for each of the columns of interest
colnames<-names(dplyr::select(CD4, -X, -Sample, -Donor, -Stim, -TB, -SM))
for(col in colnames){#this will loop through each character string in the list
    a<-subset(CD4, CD4$Stim=="PEP") #subsets the data to only contain the unstimulated samples
    #inserts each character string into the column argument of the pvals function written above
    b<-pvals_6(a, col)
    if(dim(b)[1]!=0)
    print(kable(b, caption=col)) #will return the output for each character string in the loop
}
```

#Compass Graphs
```{r}
#CD4 Mtb Responses as measured by Compass
DF<-dplyr::filter(Compass, TB != "N" & Stim == "WCL" & celltype == "CD4")

p1<-plot_compass(DF, "FS")+
    labs(y="Functionality Score")

p2<-plot_compass(DF, "PFS")+
    labs(y="Polyfunctionality Score")

grid.arrange(p1, p2, ncol=2, top="Whole Cell Lysate Scores")

DF<-dplyr::filter(Compass, TB != "N" & Stim == "PEP" & celltype == "CD4")

p1<-plot_compass(DF, "FS")+
    labs(y="Functionality Score")

p2<-plot_compass(DF,"PFS")+
    labs(y="Polyfunctionality Score")

grid.arrange(p1, p2, ncol=2, top="Peptide Pool Scores")
```

```{r, fig.width=16, fig.height=12}
library(dplyr)
DF<-merge(Compass, dplyr::select(Donor, -X, -TB, -SM), by="Donor")
ggpairs(dplyr::filter(DF, celltype=="CD4", Stim=="WCL", TB!="N")%>%
            mutate(disease = paste(TB, SM, sep=" "))%>%
            dplyr::select(age, sex, FS, PFS, disease),
        aes(col=disease))

#Boolean Graphs
The first graph is just meant to give a snapshot of all the data at once to see if there are weird anomalous data points that we will have to examine. I did not include all the individual data points on any of the graphs because it made it look incredibly cluttered, however I did allow outliers to display.
```{r Boolean Graphs, cache=TRUE, echo=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
triple_boolean_plot(CD4, "PMA")
triple_boolean_plot(CD4, "WCL")
triple_boolean_plot(CD4, "PEP")
```

#T cell Lineage Specific Grouped Analysis
These are my standard lineage specific groups. 
```{r Grouped TH Graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("Cyt.Freq", "Th1.Freq", "TH1.2.Freq", "TH2.Freq")
### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "WCL", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="WCL",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
##### Peptide
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PEP", col)+
        labs(title=paste("Frequency of ", name, " CD4+ T cells"), 
             subtitle="PEP",
             x="",
             y=paste(name, " CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
\newpage

@@

```{r Individual Cytokines but not single producers}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)
colnames<-c("IFNg", "TNFa", "IL4", "IL13")

### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
####WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}

### PEP
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("\\.", "+", col)
    name<-gsub("Freq","", name)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```

```{r Single and Double Producers Lin Spec graphs, echo=FALSE}
library(dplyr)
library(reshape2)
library(ggplot2)
library(knitr)

colnames<-c("G_T" , "IL4_IL13" )

### PMA
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             title="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
####WCL
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}

### PEP
for(col in colnames){#this will loop through each character string in the list
    #inserts each character string into the column argument of the pvals function written above
    name<-gsub("_", "+", col)
    y<-plot_with_all_stats(CD4, "PMA", col)+
        labs(title=paste("Frequency of ", name, "+ CD4+ T cells"), 
             subtitle="PMA & Ionomycin",
             x="",
             y=paste(name, "+ CD4+ T cells (%)"))
    print(y) #will return the output for each character string in the loop
}
```
